Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-10-12 code/rmarkdown/article.Rmd
+ change github username

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Hugh P. Possingham
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  All authors developed the study and drafted the manuscript. JOH performed all analyses.
lead_author_surname: "Hanson"
abstract: |
  Protected areas buffer species from anthropogenic threats and provide places for processes to continue. However, evolutionary processes are not often explicitly considered when deciding where to site protected areas because genetic data require considerable resources to obtain and analyse. Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning. We obtained occurrence and genetic data from the IntraBioDiv project for 27 species collected over the European Alps using a gridded sampling scheme. For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells. We then used the cells as planning units to generate prioritisations. Our first analysis verified that the spatial patterns of environmental and geographic variation correlated with those adaptive and neutral genetic variation (respectively). Our second analysis showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions. Our third analysis found that solutions generated using additional surrogate-based targets secured more neutral genetic variation. We discovered that solutions generated using traditional amount-based targets tend to secure most species' adaptive genetic variation. Overall, our work provides a rigorous examination of two potential surrogates for genetic variation.
significance: |
  Protected areas are used to preserve biodiversity patterns and processes. Since the resources available for conservation are limited, protected areas need to be sited in cost-effective locations. However, despite their importance, evolutionary processes are not often used to guide reserve selection. This is primarily because genetic data require extensive resources to obtain and process. Here we show that freely available environmental and geographic variables can be used as surrogates for genetic data in conservation planning.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Biodiversity and Conservation Science.
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a). This data set has previously been used to explore patterns of adaptive (eg. [@r475; @r455]) and neutral genetic variation (eg. [@r478]), and to show that species richness is not an effective surrogate for genetic diversity in conservation planning ([@r486]). It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell. We used this data set because it comprised of comparable genetic data for a range of species with different evolutionary and life histories collected using a standardised sampling scheme, and pertained to a study area with substantial environmental variation to test the effectiveness of environmental variables.

>>>>

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Hugh P. Possingham
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  All authors developed the study and drafted the manuscript. JOH performed all analyses.
lead_author_surname: "Hanson"
abstract: |
  Protected areas buffer species from anthropogenic threats and provide places for processes to continue. However, evolutionary processes are not often explicitly considered when deciding where to site protected areas because genetic data require considerable resources to obtain and analyse. Here we show that environmental and geographic distance variables are effective surrogates for adaptive and neutral intra-specific variation for conservation planning. We obtained occurrence and genetic data from the IntraBioDiv project for 27 species collected over the European Alps using a gridded sampling scheme. For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells. We then used the cells as planning units to generate prioritisations. Our first analysis verified that the spatial patterns of environmental and geographic variation correlated with those adaptive and neutral genetic variation (respectively). Our second analysis showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions. Our third analysis found that solutions generated using additional surrogate-based targets secured more neutral genetic variation. We discovered that solutions generated using traditional amount-based targets tend to secure most species' adaptive genetic variation. Overall, our work provides a rigorous examination of two potential surrogates for genetic variation.
significance: |
  Protected areas are used to preserve biodiversity patterns and processes. Since the resources available for conservation are limited, protected areas need to be sited in cost-effective locations. However, despite their importance, evolutionary processes are not often used to guide reserve selection. This is primarily because genetic data require extensive resources to obtain and process. Here we show that freely available environmental and geographic variables can be used as surrogates for genetic data in conservation planning.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a). All data, code, and results are stored in an online repository (\url{www.github.com/jeffreyhanson/genetic.surrogates}) to replicate and validate this study. This data set has previously been used to explore patterns of adaptive (eg. [@r475; @r455]) and neutral genetic variation (eg. [@r478]), and to show that species richness is not an effective surrogate for genetic diversity in conservation planning ([@r486]). It was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell. We used this data set because it comprised of comparable genetic data for a range of species with different evolutionary and life histories collected using a standardised sampling scheme, and pertained to a study area with substantial environmental variation to test the effectiveness of environmental variables.

### 2016-10-12 code/rmarkdown/supporting_information.Rmd
+ change github username

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble.tex, si-preamble.tex]
---
# Supporting Information
All data and code is stored in an online repository (\url{www.github.com/paleo13/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.

>>>>

---
fontsize: 11pt
documentclass: article
bibliography: references.bib
csl: reference-style.csl
output:
  rmarkdown::pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: [preamble.tex, si-preamble.tex]
---
# Supporting Information
All data and code is stored in an online repository (\url{www.github.com/jeffreyhanson/genetic.surrogates}) to replicate analyses in this study. See the \textit{README} file for instructions.
